hbm_paper_results

### 2024-05-01 paper_lbcc_results.Rmd
figure formatting

mean_cents_bxplt <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
  labs(y="Mean Centile", x = "Study")
```

@@

df_gam_feat_lm <- df_gam_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

@@

all_subj_means.df %>%
  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")
```

@@

control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting

@@

sample_boxplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=age_years, x=study, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
  labs(color="Sex", y = "Age (Years) \n ")

#overall sample
full_boxplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "Full Sample") %>%
  ggplot(aes(x=xlab, y=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x=" ")

sample_barplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  scale_y_continuous(labels = scales::percent) +
  #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
  labs(fill="Sex", x = "Study", y="Sex")

full_barplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "       Full Sample ") %>%
  ggplot() +
  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
  labs(color="Sex", y = "Age (Years)", y="")

ggarrange(sample_boxplot2, full_boxplot2,
          sample_barplot2, full_barplot2,
          nrow=2, ncol=2, 
          align = "hv", 
          common.legend = TRUE,
          legend = "right",
          widths=c(10, 1),
          heights=c(2, 1)
          )
```

```{r}
# study demographics
fig5_vert <- ggarrange(
  ggarrange(blank,
ggarrange(sample_boxplot2, full_boxplot2,
          sample_barplot2, full_barplot2,
          nrow=2, ncol=2, 
          #align = "v", 
          common.legend = TRUE,
          legend = "right",
          widths=c(18, 1),
          heights=c(1, 1)
          ),
blank,
nrow=1, ncol=3,
common.legend = FALSE,
widths= c(1, 20, 1)),
# boxplots
mean_cents_bxplt,
common.legend=FALSE,
nrow=2, ncol=1,
heights=c(1, 1.5),
labels="AUTO"
)

@@

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = study_mean_age), data = pred_df,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
  theme(legend.position="none") +
  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
```

>>>>

mean_cents_bxplt <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
  labs(y="Mean Centile", x = "Study")

mean_cents_bxplt_nolabel <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
  labs(y="Mean Centile", x = "Study")
```

@@

df_gam_feat_lm <- df_gam_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))

df_gam_feat_lm %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gam_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

@@

all_subj_means.df %>%
  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")
```
all assoc with study n

@@

control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.

@@

sample_boxplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=age_years, x=study, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
  labs(color="Sex", y = "Age (Years)")

#overall sample
full_boxplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "Full Sample ") %>%
  ggplot(aes(x=xlab, y=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x=" ")

sample_barplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
  #scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  scale_y_continuous(labels = scales::percent) +
  #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
  labs(fill="Sex", x = "Study", y="Percent Sex")

full_barplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "  Full Sample  ") %>%
  ggplot() +
  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
  labs(color="Sex", y = "Age (Years)", y="")
```

```{r}
# study demographics
fig5_vert <- ggarrange(
  #ggarrange(blank,
ggarrange(sample_boxplot2, full_boxplot2,
          sample_barplot2, full_barplot2,
          nrow=2, ncol=2, 
          #align = "v", 
          common.legend = TRUE,
          legend = "right",
          widths=c(18, 1),
          heights=c(1, 1)
          ),
# blank,
# nrow=1, ncol=3,
# common.legend = FALSE,
# widths= c(1, 20, 1)),
# boxplots
mean_cents_bxplt_nolabel,
common.legend=FALSE,
nrow=2, ncol=1,
heights=c(1, 1.5),
labels="AUTO"
)

@@

mean_delta_fig <- mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = study_mean_age), data = pred_df,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
  #theme(legend.position="none") +
  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")

mean_cent_deltas_df_study %>%
  dplyr::filter(abs_mean_delta_cent > 4) %>%
  dplyr::select(study, study_mean_age)

mean_cent_deltas_df_study%>%
  dplyr::filter(study == "Cornell_C2") %>%
  nrow()
```

Cornell_C2 has very low n=5 -> big differences in estimation
